Specific taxa differentiate proximal and distal microbiota in the unprepped

### 2017-07-02 submission/manuscript.Rmd
restructured intro

Colorectal cancer (CRC) is the second-leading cause of cancer-related deaths in the United States. CRC tumors vary in structure, size, morbidity and symptomology depending upon their geographical location in the colon. CRC has recently been described to fall into four molecular subtypes (@Guinney2015). Of these subtypes, tumors that arise on the distal side of the colon are infiltrating lesions that present with painful symptoms. In contrast, 47% of CRCs are caused by proximal-sided colon tumors that are bulky and project into the lumen, often remaining asymptomatic until advanced carcinogenesis (@Benedix2010). Due to the absence of symptoms, proximal-sided tumors have a significantly higher mortality rate (@Benedix2010). The distal and proximal sides of the colon differ in the amount of inflammation present and the genomic instability of precancerous cells, respectively, as well as oxygen, pH and the presence of antimicrobial peptides (@Yamauchi2012, @Albenberg2014, @Donaldson2015). Microenvironments differ not only longitutinally along the colon, but latitudinally from the epithelium to mucosa to intestinal lumen, offering several sites for different microbial communities to flourish. Given this varied physiology of the proximal-distal axis of the colon, symbiotic microbes and their metabolites likely vary between sites.

Several recent findings have shown that development and progression of CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012). Our group and others have found that the stool microbiome of patients with CRC is distinct from that of healthy people (@Zackular2014). Further studies manipulating the gut microbiome using antibiotics or other chemoagents in mice has shown that dysbiosis preceded and accelerated the development of CRC tumors (@Zackular2013). Comparison of the specific bacteria present on CRC tumors with those found on nearby healthy tissue has also identified specific bacterial species that are tumor-associated (@Dejea2014). These species include the oral pathogens _Fusobacterium nucleatum_ and _Porphyromonas asacharolytica_. Interestingly, these periodontal pathogens have been highly predictive of whether a patient had CRC tumors or not in our classification studies (@Baxter2016).  Because these studies were performed in mice or examined only shed human stool, they were unable to analyze paried samples from the proximal and distal sides of the colon. Similarly, comparisons of on- or off-tumor bacteria rarely have matched tissue from the other side of the colon from the same patient, limiting what conclusions can be drawn about the colonic microbiome overall, let alone at that specific site. Due to these limitations, the contribution of the gut microbiota to these subtypes is largely undefined. Characterizing these communities could provide needed insight into CRC etiology, including how the disruption of the healthy community could promote the initiation or proliferation of the distinct distal and proximal CRC tumors.

@@

>>>>

The human colon is an ecosystem comprised of several different microenvironments inhabited by resident bacterial members of the microbiome. Concentrations of oxygen, water and anti-microbial peptides change along the gut axis and influence what populations of microbes reside in each location. Microenvironments differ not only longitutinally along the colon, but latitudinally from the epithelium to mucosa to intestinal lumen, offering several sites for different microbial communities to flourish. The identity of these specific microbes and communities are important for understanding the etiology of complex colon diseases such as Inflammatory Bowel Disease (IBD) and Colorectal Cancer (CRC). IBD and CRC are known to be preceded or accelerated by perturbations in gut microbes (@Yamauchi2012). Further, the severity, symptoms, morbidity and mortality of these diseases is known to vary based upon the biogeographical location in which they occur. For instance, CRC tumors that arise on the distal side of the colon are infiltrating lesions that present with painful symptoms. In contrast, 47% of CRCs are caused by proximal-sided colon tumors that are sessile and form along the wall of the colon, often remaining asymptomatic until advanced carcinogenesis (@Benedix2010). The distal and proximal sides of the colon differ in the amount of inflammation present and the genomic instability of precancerous cells, respectively, in addition to variation in the previously mentioned chemical gradients (@Yamauchi2012, @Albenberg2014, @Donaldson2015). In IBD patients, disease flares in the distal colon are usually indicative of ulcerative colitis (UC), whereas Crohn's disease (CD) patients typically present with disease in the small intestine, ileum and proximal colon (cite). UC presents as large and highly inflammed mucosal ulcers, where as Crohn's disease lesions are often smaller and have areas of normal tissue distributed amongst the flare (cite). Thus, given the varied physiology of the proximal-distal axis of the colon and known differences in disease patterns at these sites, symbiotic microbes and their metabolites likely vary as well, and may influence the heterogenous disease prognoses of IBD and CRC at these locations.

Several recent findings have shown that development and progression of CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012). Our group and others have found that the stool microbiome of patients with CRC is distinct from that of healthy people (@Zackular2014). Comparison of the specific bacteria present on CRC tumors with those found on nearby healthy tissue has also identified specific bacterial species that are tumor-associated (@Dejea2014). These species include the oral pathogens _Fusobacterium nucleatum_ and _Porphyromonas asacharolytica_. Interestingly, these periodontal pathogens have been highly predictive of whether a patient had CRC tumors or not in our classification studies (@Baxter2016).  Because these studies were performed in mice or examined only shed human stool, they were unable to analyze paried samples from the proximal and distal sides of the colon. Similarly, comparisons of on- or off-tumor bacteria rarely have matched tissue from the other side of the colon from the same patient, limiting what conclusions can be drawn about the colonic microbiome overall, let alone at that specific site. Due to these limitations, the contribution of the gut microbiota to these subtypes is largely undefined. Characterizing these communities could provide needed insight into CRC etiology, including how the disruption of the healthy community could promote the initiation or proliferation of the distinct distal and proximal CRC tumors.

@@

Previous profiles of the gut biogeography only focus on prepped bowels and dont identify rare taxa. but we know that rare taxa adn small members of the community can drive tumorigenesis and whatnot. so identifying specifically UNIQUE bugs

why are L/R M/L interesting questions? because diseases progress differently depending on where they are located.
